\subsection{Surfaces}
\BEN
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\item \Emph{Identifying Surfaces}
\BEN
\item If $ab > 0$, then $a$ and $b$ have the same sign. If they are both positive, the surface is an elliptic paraboloid that opens upward. If $a$ and $b$ are both negative, the surface is an elliptic paraboloid that opens downward. 
\item If $ab < 0$, then $a$ and $b$ have opposite signs (i.e. - one is positive, the other is negative). The surface is a hyperbolic paraboloid.
\item If $a=b = 0$, then $z=0$, which is the equation of the $xy$-plane. 
\EEN
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\item \Emph{Finding the Intersection Between Two Surfaces}\\
The sphere $x^2+y^2+z^2 =9$ and the cylinder $x^2+y^2 =4$ intersect at the values of $x$ and $y$ that are satisfied by both equations. We could find these values by substituting the equation for the cylinder into the equation for the sphere:
\begin{align*}
  9 
  &= x^2+y^2+z^2 \\
  &= 4+z^2 \\
  z&=\pm\sqrt{5}
\end{align*}
Thus, the two curves intersect at $z=\sqrt{5}$ and at $z=-\sqrt{5}$. Using the equation of the sphere, we can obtain an equation for the curves that describe the intersections.
\begin{align*}
  9 
  &= x^2+y^2+(\pm\sqrt{5})^2 \\
  &= x^2+y^2+5 \\
  4&= x^2+y^2 
\end{align*}
The intersections at $z=\pm\sqrt{5}$ are described by the circle $x^2+y^2 =4$. 
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\item
\Emph{Finding the Trace of a Surface}\\
The $xy$-plane is described by the equation $z=0$. Substituting $z=0$ into the equation for the hyperbolic paraboloid gives us 
\begin{align*} 0 = \frac{x^2}{a^2} - \frac{y^2}{b^2}, \end{align*}
or simply 
\begin{align*}  y^2 = \Big(\frac{b}{a}\Big)^2x^2. \end{align*}
This equation is only satisfied by two lines: 
\begin{align*}
  y&= \frac{bx}{a}, \text{ and } y=-\frac{bx}{a}. 
\end{align*}
These two lines describe the trace of the hyperbolic parabolic in the $xy$-plane.
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\item 
\Emph{Deriving the Equation of a Surface}\\
The distance, $d_1$, from the point $(x,y,z)$ and the point $(0,1,0)$ is given by 
\begin{align*}
  d_1 = \sqrt{x^2 + (y-1)^2 + z^2}.
\end{align*}
The distance, $d_2$, from the point $(x,y,z)$ and the given plane is  
\begin{align*}
  d_2 = \big| z - 1 \big|.
\end{align*}
The points that are equidistant from the point $(0,1,0)$ and the given plane are the points that such that $d_1 = d_2$. Therefore, 
\begin{align*}
  d_1 &= d_2 \\
   \sqrt{x^2 + (y-1)^2 + z^2} &=  \big| z - 1 \big| \\
   x^2 + (y-1)^2 + z^2 &=  ( z - 1)^2 \\   
   x^2 + (y-1)^2 + z^2 &=  z^2 - 2z + 1 \\   
   x^2 + (y-1)^2  &= - 2z + 1 \\   
   z&=- \frac{1}{2} \Big(x^2 + (y-1)^2 - 1\Big)  \\   
\end{align*}
The surface is an elliptic paraboloid that opens downward and whose vertex is located at the point $(0,1,1/2)$. 

\EEN
